---
title: "P8105 HW 6"
author: "Veerapetch Petchger"
date: "2025-12-01"
output: github_document
---
# Problem 1
**The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository here. You can read their accompanying article here.**

**Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.**
```{r problem1_setup, include = FALSE}
library(tidyverse)
library(modelr)
library(purrr)
```

```{r homicide_cleaning, message = FALSE, warning = FALSE}
homicide_df = read_csv("homicide-data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    solved = if_else(disposition == "Closed by arrest", 1, 0),
    victim_age = as.numeric(victim_age)
  ) %>%
  filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
         victim_race %in% c("White", "Black")) %>% 
  select(-city, -state)
```

**For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.**
```{r baltimore_glm}
baltimore_glm =  
  homicide_df %>% 
  filter(city_state == "Baltimore, MD") %>%
  glm(
    solved ~ victim_age + victim_sex + victim_race,
    data = ., family = binomial()
  )

baltimore_glm  %>% 
  broom::tidy(conf.int = TRUE, exponentiate = TRUE) %>%
  filter(term == "victim_sexMale") %>%
  select(term, estimate, conf.low, conf.high)
```
**Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.**
```{r city_glm, warning = FALSE}
city_glm =
  homicide_df %>% 
  group_by(city_state) %>% 
  nest() %>% 
  mutate(
    fit = map(data, ~ glm(solved ~ victim_age + victim_sex + victim_race,
                          data = ., family = binomial())),
    tidy_fit = map(.x = fit, ~broom::tidy(.x, conf.int = TRUE, exponentiate = TRUE) )
  ) %>% 
  select(city_state, tidy_fit) %>% 
  unnest(tidy_fit) %>% 
  filter(term == "victim_sexMale") %>%
  select(city_state, estimate, conf.low, conf.high)
city_glm
```
**Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.**
```{r city_glm_plot}
city_glm %>% 
  ggplot(aes(x = reorder(city_state, estimate), y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  coord_flip() +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 7)) +
  labs(
    title = "Adjusted odds ratio of solved homicides between male and female victims",
    x = "City",
    y = "Adjusted odds ratio (male vs female victims)"
  )
```
The majority of odds ratios are less than one, suggesting that cases female homicide victims in these cities are less often resolved. Of these cities, the odds ratio is statistically significant, implying sufficient evidence of a true difference in case disposition across sexes. Three cities report an odds ratio above 1, but they are not statistically significant since the lower bound of the interval is below 1.

# Problem 2
**For this problem, we’ll use the Central Park weather data we’ve seen elsewhere. The code chunk below will import these data from the p8105.datasets package.**
```{r problem2_setup, include = FALSE}
library(p8105.datasets)
data("weather_df")
```
**The bootstrap is helpful when you’d like to perform inference for a parameter / value / summary that doesn’t have an easy-to-write-down distribution in the usual repeated sampling framework. We’ll focus on a simple linear regression with tmax as the response with tmin and prcp as the predictors, and are interested in the distribution of two quantities estimated from these data:**

- 𝑟̂ 2

- 𝛽̂ 1𝛽̂ 2

**Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution of your estimates, and describe these in words. Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for 𝑟̂ 2 and 𝛽̂ 1𝛽̂ 2.**

**Note: broom::glance() is helpful for extracting 𝑟̂ 2 from a fitted regression, and broom::tidy() (with some additional wrangling) should help in computing 𝛽̂ 1𝛽̂ 2.**
```{r weather_boot}
set.seed(123)
weather_boot = 
  weather_df %>% 
  drop_na(tmax, tmin, prcp) %>%
  bootstrap(n = 5000, id = "strap_id") %>% 
  mutate(
    fit = map(strap, ~lm(tmax ~ tmin + prcp, data = .x)),
    glance = map(fit, broom::glance),
    tidy = map(fit, broom::tidy)
  ) %>% 
  mutate(
    r2_weather = map_dbl(glance, "r.squared"),
    beta_ratio = map_dbl(tidy, ~ {
      b1 = .x %>% 
        filter(term == "tmin") %>% pull(estimate)
      b2 = .x %>% 
        filter(term == "prcp") %>% pull(estimate)
      b1 / b2
    })
  ) %>% 
  select(strap_id, r2_weather, beta_ratio)

weather_long =
  weather_boot %>% 
  pivot_longer(
    cols = c("r2_weather", "beta_ratio"),
    names_to = "parameter",
    values_to = "estimate"
  )

weather_long %>% 
  ggplot(aes(x = estimate)) + 
  geom_histogram(bins = 50) +
  facet_wrap(. ~ parameter, scales = "free") +
  labs(
    title = "Bootstrap distributions for r^2 and β1/β2",
    x = "Bootstrap estimate",
    y = "Count"
  ) +
  theme_minimal()
```

```{r weather_ci}
weather_long %>% 
  group_by(parameter) %>% 
  summarize(
    ci_lower = quantile(estimate, 0.275),
    ci_upper = quantile(estimate, 0.975)
  )
```

# Problem 3
**Load and clean the data for regression analysis (i.e. use appropriate variable names, convert numeric to factor where appropriate, check for the presence of missing data, etc.).**
```{r birthweight_clean}
birthweight_df = 
  read_csv("birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    babysex = factor(babysex,
                     levels = c(1,2),
                     labels = c("male", "female")),
    frace = factor(frace,
                   levels = c(1, 2, 3, 4, 8, 9),
                   labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown" )),
    malform = factor(malform,
                     levels = c(0,1),
                     labels = c("absent", "present")),
    mrace = factor(mrace,
                   levels = c(1, 2, 3, 4, 8, 9),
                   labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown" ))
  )
```

**Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.**
```{r proposed_model}
birthweight_mod = 
  birthweight_df %>% 
  lm(bwt ~ babysex + gaweeks + malform + mheight + momage + mrace + parity + pnumlbw + pnumsga + smoken, data = .)
birthweight_mod %>% 
  summary()
```

```{r proposed_model_plot}
birthweight_df %>% 
  add_predictions(birthweight_mod) %>% 
  add_residuals(birthweight_mod) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = "Fitted values vs Residuals of Proposed Birthweight Model",
    x = "Fitted values",
    y = "Residuals"
  ) +
  theme_minimal()
```
I created a multiple linear regression model with respect to birthweight (`bwt`). The model investigates the relationships between several variables that I felt are influential to determining `bwt`. These variables are the baby's sex (`babysex`), gestational age in weeks (`gaweeks`), the presence of malformations (`malform`), the mother's race (`mrace`), number of prior live births (`parity`), number of prior low birth weight babies (`pnumlbw`), number of prior small for gestational age babies (`pnumsga`), and the average number of cigarettes smoked per day (`smoken`). Measurements related to the babies' phenotypes such as their head circumference (`bhead`) and length (`blength`) at birth were excluded as they are competing outcomes with `bwt`.

**Compare your model to two others:**

- One using length at birth and gestational age as predictors (main effects only)
```{r}
bwt_model1 = 
  birthweight_df %>% 
  lm(bwt ~ blength + gaweeks, data = .)
```

- One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
```{r}
bwt_model2 = 
  birthweight_df %>% 
  lm(bwt ~ bhead * blength * babysex, data = .)
```

**Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.**

**Note that although we expect your model to be reasonable, model building itself is not a main idea of the course and we don’t necessarily expect your model to be “optimal”.**

